---
title: "BDHS EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(DTDA)
```

```{r}
BDHS_data <- read.csv("./BDHS_data_1.csv")
gen <- 1900 + floor(BDHS_data$dateOfBirth / 12) 
gen <- round(gen - 5, -1)
BDHS_data <- cbind(BDHS_data, gen)
df <- BDHS_data[BDHS_data$delta1 == 1, ]
```

**Age at First Birth (overall)** <br>
```{r}
summary(df$ageAtFirstBirth)
sd(df$ageAtFirstBirth)
```

```{r}
boxplot(df$ageAtFirstBirth,
        main = "Boxplot of Overall Age at First Birth")
```

**Number of Children (overall)** <br>
```{r}
summary(df$birthCount)
sd(df$birthCount)
```

```{r}
boxplot(df$birthCount,
        main = "Boxplot of Overall Number of Children")
```

**Age at First Birth grouped by division** <br>
```{r}
df %>% group_by(division) %>%
  summarise(min = min(ageAtFirstBirth),
            firstQu = quantile(ageAtFirstBirth, 0.25),
            average =  mean(ageAtFirstBirth),
            median = median(ageAtFirstBirth),
            thirdQu = quantile(ageAtFirstBirth, 0.75),
            max = max(ageAtFirstBirth),
            std = sd(ageAtFirstBirth))
```

**Number of Children grouped by division** <br>
```{r}
df %>% group_by(division) %>%
  summarise(min = min(birthCount),
            firstQu = quantile(birthCount, 0.25),
            average =  mean(birthCount),
            median = median(birthCount),
            thirdQu = quantile(birthCount, 0.75),
            max = max(birthCount),
            std = sd(birthCount))
```

```{r}
par(mfrow = c(1, 2))
boxplot(df$birthCount ~ df$division,
        main = "Number of Children")
boxplot(df$ageAtFirstBirth ~ df$division,
        main = "Age at First Birth")
```
This result is the same with Hsieh's Thesis. <br>

**Age at First Birth grouped by generation** <br>
```{r}
df %>% group_by(gen) %>%
  summarise(min = min(ageAtFirstBirth),
            firstQu = quantile(ageAtFirstBirth, 0.25),
            average =  mean(ageAtFirstBirth),
            median = median(ageAtFirstBirth),
            thirdQu = quantile(ageAtFirstBirth, 0.75),
            max = max(ageAtFirstBirth),
            std = sd(ageAtFirstBirth))
```

**Number of Children grouped by generation** <br>
```{r}
df %>% group_by(gen) %>%
  summarise(min = min(birthCount),
            firstQu = quantile(birthCount, 0.25),
            average =  mean(birthCount),
            median = median(birthCount),
            thirdQu = quantile(birthCount, 0.75),
            max = max(birthCount),
            std = sd(birthCount))
```

```{r}
par(mfrow = c(1, 2))
boxplot(df$birthCount ~ df$gen,
        main = "Number of Children")
boxplot(df$ageAtFirstBirth ~ df$gen,
        main = "Age at First Birth")
```

```{r}
par(mfrow = c(1, 2))
boxplot(df$ageAtFirstBirth ~ df$division,
        main = "Age at First Birth by division")
boxplot(df$ageAtFirstBirth ~ df$gen,
        main = "Age at First Birth by generation")
```

```{r}
par(mfrow = c(1, 2))
boxplot(df$birthCount ~ df$division,
        main = "Number of Children by division")
boxplot(df$birthCount ~ df$gen,
        main = "Number of Children by generation")
```
The younger generation is truncated from the dataset.

**Boundary Condition** <br>
```{r}
min(df$Y1)
min(df$C)
max(df$Y1)
max(df$C)
```

```{r}
min(BDHS_data$Y1)
min(BDHS_data$C)
max(BDHS_data$Y1)
max(BDHS_data$C)
```


```{r}
df %>% group_by(gen) %>%
  summarise(a_y1 = min(Y1),
            a_C = min(C),
            b_y1 = max(Y1),
            b_C = max(C))
```

```{r}
df %>% group_by(division) %>%
  summarise(a_y1 = min(Y1),
            a_C = min(C),
            b_y1 = max(Y1),
            b_C = max(C))
```

**truncation probability (overall)** <br>
```{r}
trunc_prob_func(df$Y1, df$C, unique(df$m))
```

**truncation probability (grouped by division)** <br>
```{r}
df %>%
  group_by(division) %>%
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))
```

**truncation probability (grouped by generation)** <br>
```{r}
df %>%
  group_by(gen) %>%
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))
```


```{r}
## division 1
df %>% 
  filter(division == 1) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 2
df %>% 
  filter(division == 2) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 3
df %>% 
  filter(division == 3) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 4
df %>% 
  filter(division == 4) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 5
df %>% 
  filter(division == 5) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 6
df %>% 
  filter(division == 6) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

```

```{r}
nrow(df[df$gen == 1990, ])
```

**Censoring rate (overall)** <br>
```{r}
mean(df$delta2 == 0)
```

**Censoring rate grouped by division** <br>
```{r}
df %>% group_by(division) %>%
  summarise(censoring.rate = mean(delta2 == 0))
```

**Censoring rate grouped by generation** <br>
```{r}
df %>% group_by(gen) %>%
  summarise(censoring.rate = mean(delta2 == 0))
```

```{r}
## division 1
df %>% 
  filter(division == 1) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 2
df %>% 
  filter(division == 2) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 3
df %>% 
  filter(division == 3) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 4
df %>% 
  filter(division == 4) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 5
df %>% 
  filter(division == 5) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 6
df %>% 
  filter(division == 6) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

```

**Marginal distribution of Y1** <br>
```{r}
Sy1 <- lynden(X = df$Y1, V = df$C, boot = FALSE)
plot(Sy1$time, Sy1$survival, type = "l")
```

```{r}
for(div in 1:6) {
  df.div <- df[df$division == div, ]
  Sy1.div <- lynden(X = df.div$Y1, V = df.div$C, boot = FALSE)
  
  if(div == 1) {
    plot(Sy1$time, Sy1$survival, type = "l", xlim = c(0, 200))
    points(Sy1.div$time, Sy1.div$survival, type = "l", lty = div + 1, col = div + 1)
  } else {
    points(Sy1.div$time, Sy1.div$survival, type = "l", lty = div + 1, col = div + 1)
  }
}

legend("topright",
       c("Overall", "Barisal", "Chittagong", "Dhaka", "Khuna", "Rajshahi", "Sylhet"),
       lty = 1:7,
       col = 1:7)
```

```{r}
gen <- c(1950, 1960, 1970, 1980)

for(i in 1:4) {
  df.gen <- df[df$gen == gen[i], ]
  Sy1.gen <- lynden(X = df.gen$Y1, V = df.gen$C, boot = FALSE)
  
  if(i == 1) {
    plot(Sy1$time, Sy1$survival, type = "l", xlim = c(0, 200))
    points(Sy1.gen$time, Sy1.gen$survival, type = "l", lty = i + 1, col = i + 1)
  } else {
    points(Sy1.gen$time, Sy1.gen$survival, type = "l", lty = i + 1, col = i + 1)
  }
}

legend("topright",
       c("Overall", gen),
       lty = 1:5,
       col = 1:5)
```

```{r}
gen <- c(1950, 1960, 1970, 1980)
div <- c("Barisal", "Chittagong", "Dhaka", "Khuna", "Rajshahi", "Sylhet")

for(j in 1:6) {
  for(i in 1:4) {
    df.div.gen <- df[df$gen == gen[i] & df$division == j, ]
    Sy1.div.gen <- lynden(X = df.div.gen$Y1, V = df.div.gen$C, boot = FALSE)
    
    if(i == 1) {
      plot(Sy1.div.gen$time,
           Sy1.div.gen$survival,
           type = "l",
           lty = i + 1,
           col = i + 1,
           xlim = c(0, 200),
           main = paste("Survival of Y1 in", div[j]))
    } else {
      points(Sy1.div.gen$time, Sy1.div.gen$survival, type = "l", lty = i + 1, col = i + 1)
    }
  }
  
  legend("topright", legend = gen, lty = 2:5, col = 2:5)
}
```

```{r}
gen <- c(1950, 1960, 1970, 1980)
div <- c("Barisal", "Chittagong", "Dhaka", "Khuna", "Rajshahi", "Sylhet")

for(i in 1:4) {
  for(j in 1:6) {
    df.gen.div <- df[df$gen == gen[i] & df$division == j, ]
    Sy1.gen.div <- lynden(X = df.gen.div$Y1, V = df.gen.div$C, boot = FALSE)
    
    if(j == 1) {
      plot(Sy1.gen.div$time,
           Sy1.gen.div$survival,
           type = "l",
           lty = j + 1,
           col = j + 1,
           xlim = c(0, 200),
           main = paste("Survival of Y1 in", gen[i]))
    } else {
      points(Sy1.gen.div$time, Sy1.gen.div$survival, type = "l", lty = j + 1, col = j + 1)
    }
  }
  
  legend("topright", legend = div, lty = 2:7, col = 2:7)
}
```


**Truncation Probability** <br>
```{r}
Sc <- lynden(X = df$C, U = df$Y1, boot = FALSE)
mass.c <- c(Sc$cumulative.df[1], diff(Sc$cumulative.df))
Fy1_func <- stepfun(x = Sy1$time,
                    y = c(Sy1$cumulative.df, 1),
                    right = TRUE,
                    f = 1)
trunc.prob<- sum(Fy1_func(Sc$time) * mass.c)
```

```{r}
trunc.prob.div <- numeric(length = 6)
for(i in 1:6) {
  df.div <- df[df$division == i, ]
  Sy1.div <- lynden(X = df.div$Y1, V = df.div$C, boot = FALSE)
  Sc.div <- lynden(X = df.div$C, U = df.div$Y1, boot = FALSE)
  mass.c.div <- c(Sc.div$cumulative.df[1], diff(Sc.div$cumulative.df))
  Fy1.div_func <- stepfun(x = Sy1.div$time,
                          y = c(Sy1.div$cumulative.df, 1),
                          right = TRUE,
                          f = 1)
  trunc.prob.div[i] <- sum(Fy1.div_func(Sc.div$time) * mass.c.div)
}

trunc.prob.div
```

```{r}
trunc.prob.gen <- numeric(length = 4)
gen <- c(1950, 1960, 1970, 1980)

for(i in 1:4) {
  df.gen <- df[df$gen == gen[i], ]
  Sy1.gen <- lynden(X = df.gen$Y1, V = df.gen$C, boot = FALSE)
  Sc.gen <- lynden(X = df.gen$C, U = df.gen$Y1, boot = FALSE)
  mass.c.gen <- c(Sc.gen$cumulative.df[1], diff(Sc.gen$cumulative.df))
  Fy1.gen_func <- stepfun(x = Sy1.gen$time,
                          y = c(Sy1.gen$cumulative.df, 1),
                          right = TRUE,
                          f = 1)
  trunc.prob.gen[i] <- sum(Fy1.gen_func(Sc.gen$time) * mass.c.gen)
}

trunc.prob.gen
```

```{r}
trunc.prob.div.gen <- matrix(nrow = 6, ncol = 4)
gen <- c(1950, 1960, 1970, 1980)

for(i in 1:6) {
  for(j in 1:4) {
    df.div.gen <- df[df$gen == gen[j] & df$division == i, ]
    Sy1.div.gen <- lynden(X = df.div.gen$Y1, V = df.div.gen$C, boot = FALSE)
    Sc.div.gen <- lynden(X = df.div.gen$C, U = df.div.gen$Y1, boot = FALSE)
    mass.c.div.gen <- c(Sc.div.gen$cumulative.df[1], diff(Sc.div.gen$cumulative.df))
    Fy1.div.gen_func <- stepfun(x = Sy1.div.gen$time,
                                y = c(Sy1.div.gen$cumulative.df, 1),
                                right = TRUE,
                                f = 1)
    trunc.prob.div.gen[i, j] <- sum(Fy1.div.gen_func(Sc.div.gen$time) * mass.c.div.gen)
  }
}

trunc.prob.div.gen
```








