---
title: "BDHS EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(DTDA)
```

```{r}
BDHS_data <- read.csv("./BDHS_data_1.csv")
gen <- 1900 + floor(BDHS_data$dateOfBirth / 12) 
gen <- round(gen - 5, -1)
BDHS_data <- cbind(BDHS_data, gen)
df <- BDHS_data[BDHS_data$delta1 == 1, ]
```

**Age at First Birth (overall)** <br>
```{r}
summary(df$ageAtFirstBirth)
sd(df$ageAtFirstBirth)
```

```{r}
boxplot(df$ageAtFirstBirth,
        main = "Boxplot of Overall Age at First Birth")
```

**Number of Children (overall)** <br>
```{r}
summary(df$birthCount)
sd(df$birthCount)
```

```{r}
boxplot(df$birthCount,
        main = "Boxplot of Overall Number of Children")
```

**Age at First Birth grouped by division** <br>
```{r}
df %>% group_by(division) %>%
  summarise(min = min(ageAtFirstBirth),
            firstQu = quantile(ageAtFirstBirth, 0.25),
            average =  mean(ageAtFirstBirth),
            median = median(ageAtFirstBirth),
            thirdQu = quantile(ageAtFirstBirth, 0.75),
            max = max(ageAtFirstBirth),
            std = sd(ageAtFirstBirth))
```

**Number of Children grouped by division** <br>
```{r}
df %>% group_by(division) %>%
  summarise(min = min(birthCount),
            firstQu = quantile(birthCount, 0.25),
            average =  mean(birthCount),
            median = median(birthCount),
            thirdQu = quantile(birthCount, 0.75),
            max = max(birthCount),
            std = sd(birthCount))
```

```{r}
par(mfrow = c(1, 2))
boxplot(df$birthCount ~ df$division,
        main = "Number of Children")
boxplot(df$ageAtFirstBirth ~ df$division,
        main = "Age at First Birth")
```
This result is the same with Hsieh's Thesis. <br>

**Age at First Birth grouped by generation** <br>
```{r}
df %>% group_by(gen) %>%
  summarise(min = min(ageAtFirstBirth),
            firstQu = quantile(ageAtFirstBirth, 0.25),
            average =  mean(ageAtFirstBirth),
            median = median(ageAtFirstBirth),
            thirdQu = quantile(ageAtFirstBirth, 0.75),
            max = max(ageAtFirstBirth),
            std = sd(ageAtFirstBirth))
```

**Number of Children grouped by generation** <br>
```{r}
df %>% group_by(gen) %>%
  summarise(min = min(birthCount),
            firstQu = quantile(birthCount, 0.25),
            average =  mean(birthCount),
            median = median(birthCount),
            thirdQu = quantile(birthCount, 0.75),
            max = max(birthCount),
            std = sd(birthCount))
```

```{r}
par(mfrow = c(1, 2))
boxplot(df$birthCount ~ df$gen,
        main = "Number of Children")
boxplot(df$ageAtFirstBirth ~ df$gen,
        main = "Age at First Birth")
```

```{r}
par(mfrow = c(1, 2))
boxplot(df$ageAtFirstBirth ~ df$division,
        main = "Age at First Birth by division")
boxplot(df$ageAtFirstBirth ~ df$gen,
        main = "Age at First Birth by generation")
```

```{r}
par(mfrow = c(1, 2))
boxplot(df$birthCount ~ df$division,
        main = "Number of Children by division")
boxplot(df$birthCount ~ df$gen,
        main = "Number of Children by generation")
```
The younger generation is truncated from the dataset.

**Boundary Condition** <br>
```{r}
min(df$Y1)
min(df$C)
max(df$Y1)
max(df$C)
```

```{r}
min(BDHS_data$Y1)
min(BDHS_data$C)
max(BDHS_data$Y1)
max(BDHS_data$C)
```


```{r}
df %>% group_by(gen) %>%
  summarise(a_y1 = min(Y1),
            a_C = min(C),
            b_y1 = max(Y1),
            b_C = max(C))
```

```{r}
df %>% group_by(division) %>%
  summarise(a_y1 = min(Y1),
            a_C = min(C),
            b_y1 = max(Y1),
            b_C = max(C))
```

**truncation probability (overall)** <br>
```{r}
trunc_prob_func(df$Y1, df$C, unique(df$m))
```

**truncation probability (grouped by division)** <br>
```{r}
df %>%
  group_by(division) %>%
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))
```

**truncation probability (grouped by generation)** <br>
```{r}
df %>%
  group_by(gen) %>%
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))
```


```{r}
## division 1
df %>% 
  filter(division == 1) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 2
df %>% 
  filter(division == 2) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 3
df %>% 
  filter(division == 3) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 4
df %>% 
  filter(division == 4) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 5
df %>% 
  filter(division == 5) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

## division 6
df %>% 
  filter(division == 6) %>%
  group_by(gen) %>% 
  summarise(pi.hat = trunc_prob_func(Y1, C, unique(m)))

```

```{r}
nrow(df[df$gen == 1990, ])
```

**Censoring rate (overall)** <br>
```{r}
mean(df$delta2 == 0)
```

**Censoring rate grouped by division** <br>
```{r}
df %>% group_by(division) %>%
  summarise(censoring.rate = mean(delta2 == 0))
```

**Censoring rate grouped by generation** <br>
```{r}
df %>% group_by(gen) %>%
  summarise(censoring.rate = mean(delta2 == 0))
```

```{r}
## division 1
df %>% 
  filter(division == 1) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 2
df %>% 
  filter(division == 2) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 3
df %>% 
  filter(division == 3) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 4
df %>% 
  filter(division == 4) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 5
df %>% 
  filter(division == 5) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

## division 6
df %>% 
  filter(division == 6) %>%
  group_by(gen) %>% 
  summarise(censoring.rate = mean(delta2 == 0))

```

**Marginal distribution of Y1** <br>
```{r}
Sy1 <- lynden(X = df$Y1, V = df$C, display.S = FALSE)
plot(Sy1$time, Sy1$survival, type = "l")
```


